### 2014-02-28 exception-analysis/reviews/esop2014/ESOP-REBUTTAL.txt
icfp etc.

We would like to thank the reviewers for their constructive feedback.

==REMARKS==

    CBN/CBV:

        The primary reason for choosing a call-by-name language lies in our
        ambition of scaling the analysis to the full Haskell language. This
        choice does influence several technical details (the need for an
        exception type for both argument and result type of a function, as
        well as an exception annotation on the function space constructor
        itself). Somewhat counter-intuitively the imprecise exception semantics
        complicate things for the analysis as well (we need non-emptiness
        guards). But these are details we want to handle given our ambitions.

        Some of the general ideas of the algorithm, such as the use of
        implication constraints, could be transfered to an analysis for a
        call-by-value language.

    Related work:

        Due to the page limit we had to make some concessions in the Related
        Work section. We will try to incorporate some of the suggestions,
        criticism and pointers to other related work offered by the reviewers.

    Constraint solving:

        We do not have a fully worked-out proof of correctness of the
        constraint solver, but one can at least convince oneself that the
        algorithm terminates as follows:

            During each iteration of the while-loop either W will decrease or
            \theta will increase. As long as \Lambda_\iota satisfy the
            ascending chain condition (trivially if they are finite) the
            algorithm will terminate.

        To see that the algorithm finds the least solution, note that
        constraints appearing on the left-hand side of an implication are
        monotonic when used as a predicate (as in the if-then statement
        within the while-loop). Intuitively: once a guard holds in iteration
        $n$ of the algorithm, it will never fail to hold during a later
        iteration. This property can fail to hold for more expressive
        constraint languages, as explained in Section 4.5.

==QUESTIONS==

    - Reviewer 1(iv): The expressions which are typeable in the system in
            Figure 2 are exactly those expressions which are typeable in the
            underlying type system. Furthermore, types in the analysis have
            the same shape as the underlying types. So 'f 3 4 5' is indeed not
            typeable, but this is because it would not be typeable in the
            underlying (monomorphic) type system with f : A -> B for base
            types A, B. Instead, f would be given the type

                forall a b c d. a -> b -> c -> d.

            An alternative presentation in terms of an annotated type system
            (intertwining both the underlying type system and the analysis)
            would be possible, but would in our opinion complicate the
            presentation more than it would help to clarify.

    - Reviewer 1(v): Section 6 presents an extension of the analysis. It is
            not included in Figure 2.

    - Reviewer 2: Section 6 sketches an algorithm that handles polymorphism.

    - Reviewer 3: "Isn't the type language of 4.2 missing some base type
            constructors?": No, these are all subsumed by the type variable
            \alpha. Concrete base type are part of the underlying type system,
            not the analysis.