### 2015-09-01 program-analysis-implementations/NielsonNielsonAmtoft/NielsonNielsonAmtoft.tex
program-analysis-implementations

\documentclass{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[paperwidth=250cm,paperheight=250cm]{geometry}
\usepackage[cm]{fullpage}
\usepackage{stmaryrd}
\usepackage{xcolor}
\begin{document}

\begin{description}
\item[Substitution] \[\left[\begin{split}\alpha_{11}& \mapsto \dot\alpha_{222}\\\alpha_{121}& \mapsto \dot\alpha_{132}\\\alpha_{12}& \mapsto \dot\alpha_{205}\\\alpha_{145}& \mapsto \left(\dot\alpha_{181}\xrightarrow{\dot\beta_{160}}\left(\dot\alpha_{158} \times \dot\alpha_{159}\right)\right)\\\alpha_{14}& \mapsto \dot\alpha_{150}\\\alpha_{179}& \mapsto \left(\dot\alpha_{189} \times \dot\alpha_{190}\right)\\\alpha_{18}& \mapsto \dot\alpha_{131}\\\alpha_{1}& \mapsto \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\\alpha_{203}& \mapsto \dot\alpha_{213}\\\alpha_{20}& \mapsto \dot\alpha_{92}\\\alpha_{239}& \mapsto \left(\dot\alpha_{250}\xrightarrow{\dot\beta_{253}}\dot\alpha_{891}\right)\\\alpha_{23}& \mapsto \mathbf{Chan}\ \dot\alpha_{50}\\\alpha_{276}& \mapsto \left(\dot\alpha_{406}\xrightarrow{\dot\beta_{310}}\left(\left(\dot\alpha_{335}\xrightarrow{\dot\beta_{338}}\dot\alpha_{336}\right) \times \dot\alpha_{309}\right)\right)\\\alpha_{29}& \mapsto \mathbf{Chan}\ \dot\alpha_{31}\\\alpha_{36}& \mapsto \left(\dot\alpha_{72}\xrightarrow{\dot\beta_{41}}\left(\mathbf{Chan}\ \dot\alpha_{51} \times \dot\alpha_{40}\right)\right)\\\alpha_{3}& \mapsto \left(\dot\alpha_{818}\xrightarrow{\dot\beta_{568}}\dot\alpha_{567}\right)\\\alpha_{403}& \mapsto \left(\left(\dot\alpha_{451}\xrightarrow{\dot\beta_{454}}\dot\alpha_{452}\right) \times \dot\alpha_{453}\right)\\\alpha_{497}& \mapsto \dot\alpha_{541}\\\alpha_{4}& \mapsto \dot\alpha_{499}\\\alpha_{641}& \mapsto \left(\dot\alpha_{880}\xrightarrow{\dot\beta_{744}}\dot\alpha_{739}\right)\\\alpha_{642}& \mapsto \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\\\alpha_{681}& \mapsto \dot\alpha_{902}\\\alpha_{6}& \mapsto \left(\dot\alpha_{333}\xrightarrow{\dot\beta_{337}}\dot\alpha_{334}\right)\\\alpha_{70}& \mapsto \left(\mathbf{Chan}\ \dot\alpha_{803} \times \dot\alpha_{76}\right)\\\alpha_{719}& \mapsto \left(\dot\alpha_{740}\xrightarrow{\dot\beta_{745}}\dot\alpha_{741}\right)\\\alpha_{87}& \mapsto \dot\alpha_{103}\ \mathbf{Com}\ \dot\beta_{104}\\\dot\alpha_{123}& \mapsto \dot\alpha_{132}\\\dot\alpha_{126}& \mapsto \dot\alpha_{803}\\\dot\alpha_{147}& \mapsto \dot\alpha_{803}\\\dot\alpha_{157}& \mapsto \dot\alpha_{181}\\\dot\alpha_{184}& \mapsto \dot\alpha_{803}\\\dot\alpha_{208}& \mapsto \dot\alpha_{803}\\\dot\alpha_{236}& \mapsto \dot\alpha_{803}\\\dot\alpha_{240}& \mapsto \left(\dot\alpha_{250}\xrightarrow{\dot\beta_{253}}\dot\alpha_{891}\right)\\\dot\alpha_{243}& \mapsto \dot\alpha_{803}\\\dot\alpha_{248}& \mapsto \dot\alpha_{885}\\\dot\alpha_{249}& \mapsto \dot\alpha_{890}\\\dot\alpha_{251}& \mapsto \dot\alpha_{891}\\\dot\alpha_{278}& \mapsto \dot\alpha_{803}\\\dot\alpha_{281}& \mapsto \dot\alpha_{891}\\\dot\alpha_{286}& \mapsto \dot\alpha_{885}\\\dot\alpha_{291}& \mapsto \dot\alpha_{890}\\\dot\alpha_{292}& \mapsto \dot\alpha_{891}\\\dot\alpha_{307}& \mapsto \dot\alpha_{406}\\\dot\alpha_{308}& \mapsto \left(\dot\alpha_{335}\xrightarrow{\dot\beta_{338}}\dot\alpha_{336}\right)\\\dot\alpha_{38}& \mapsto \dot\alpha_{72}\\\dot\alpha_{39}& \mapsto \mathbf{Chan}\ \dot\alpha_{51}\\\dot\alpha_{405}& \mapsto \dot\alpha_{803}\\\dot\alpha_{413}& \mapsto \dot\alpha_{891}\\\dot\alpha_{422}& \mapsto \dot\alpha_{885}\\\dot\alpha_{431}& \mapsto \dot\alpha_{890}\\\dot\alpha_{432}& \mapsto \dot\alpha_{891}\\\dot\alpha_{500}& \mapsto \dot\alpha_{803}\\\dot\alpha_{507}& \mapsto \dot\alpha_{891}\\\dot\alpha_{516}& \mapsto \dot\alpha_{885}\\\dot\alpha_{525}& \mapsto \dot\alpha_{890}\\\dot\alpha_{526}& \mapsto \dot\alpha_{891}\\\dot\alpha_{566}& \mapsto \dot\alpha_{818}\\\dot\alpha_{612}& \mapsto \dot\alpha_{803}\\\dot\alpha_{615}& \mapsto \dot\alpha_{891}\\\dot\alpha_{620}& \mapsto \dot\alpha_{885}\\\dot\alpha_{625}& \mapsto \dot\alpha_{890}\\\dot\alpha_{626}& \mapsto \dot\alpha_{891}\\\dot\alpha_{677}& \mapsto \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\\\dot\alpha_{716}& \mapsto \dot\alpha_{902}\\\dot\alpha_{721}& \mapsto \left(\dot\alpha_{740}\xrightarrow{\dot\beta_{745}}\dot\alpha_{741}\right)\\\dot\alpha_{722}& \mapsto \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\\\dot\alpha_{729}& \mapsto \dot\alpha_{902}\\\dot\alpha_{738}& \mapsto \dot\alpha_{880}\\\dot\alpha_{743}& \mapsto \dot\alpha_{914}\\\dot\alpha_{75}& \mapsto \dot\alpha_{803}\\\dot\alpha_{806}& \mapsto \dot\alpha_{885}\\\dot\alpha_{809}& \mapsto \dot\alpha_{891}\\\dot\alpha_{819}& \mapsto \dot\alpha_{885}\\\dot\alpha_{820}& \mapsto \dot\alpha_{891}\\\dot\alpha_{827}& \mapsto \dot\alpha_{902}\\\dot\alpha_{834}& \mapsto \dot\alpha_{914}\\\dot\alpha_{847}& \mapsto \dot\alpha_{880}\\\dot\alpha_{848}& \mapsto \dot\alpha_{902}\\\dot\alpha_{849}& \mapsto \dot\alpha_{914}\\\dot\alpha_{877}& \mapsto \dot\alpha_{891}\\\dot\alpha_{89}& \mapsto \dot\alpha_{803}\\\dot\alpha_{93}& \mapsto \dot\alpha_{803}; \epsilon\end{split}\right]\]
\item[Type]         \[\left(\left(\dot\alpha_{818}\xrightarrow{\dot\beta_{252}}\dot\alpha_{891}\right)\xrightarrow{\beta_{876}}\left(\dot\alpha_{880}\xrightarrow{\dot\beta_{745}}\dot\alpha_{880}\right)\right)\]
\item[Behaviour]    \[\left\{\right\}\]
\item[Constraints]  \[\left\{\dot\alpha_{742}\subseteq \dot\alpha_{880}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{880}\subseteq \dot\alpha_{902}, \dot\alpha_{880}\subseteq \dot\alpha_{914}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{663}\right\}\subseteq \beta_{720}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{702}\right\}\subseteq \dot\beta_{744}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{720}\right\}\subseteq \beta_{876}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{745}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{746}\right\}\]
\end{description}
\begin{gather}
\frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \mathbf{snd}\\&  : \left(\left(\left(\dot\alpha_{818}\xrightarrow{\dot\beta_{568}}\dot\alpha_{567}\right) \times \dot\alpha_{499}\right)\xrightarrow{\beta_{5}}\dot\alpha_{499}\right)\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }\frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \mathbf{(,)}\\&  : \left(\left(\dot\alpha_{333}\xrightarrow{\dot\beta_{337}}\dot\alpha_{334}\right)\xrightarrow{\beta_{8}}\left(\alpha_{7}\xrightarrow{\beta_{9}}\left(\left(\dot\alpha_{333}\xrightarrow{\dot\beta_{337}}\dot\alpha_{334}\right) \times \alpha_{7}\right)\right)\right)\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }\frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \mathbf{True}\\&  : \mathbf{Bool}\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }\frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash f\\&  : \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }\frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \mathbf{snd}\\&  : \left(\left(\dot\alpha_{222} \times \dot\alpha_{205}\right)\xrightarrow{\beta_{13}}\dot\alpha_{205}\right)\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }\frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \mathbf{(,)}\\&  : \left(\dot\alpha_{150}\xrightarrow{\beta_{16}}\left(\alpha_{15}\xrightarrow{\beta_{17}}\left(\dot\alpha_{150} \times \alpha_{15}\right)\right)\right)\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }\frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \mathbf{sync}\\&  : \left(\dot\alpha_{131}\ \mathbf{Com}\ \beta_{19}\xrightarrow{\beta_{19}}\dot\alpha_{131}\right)\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }\frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \mathbf{send}\\&  : \left(\left(\mathbf{Chan}\ \dot\alpha_{92} \times \dot\alpha_{92}\right)\xrightarrow{\beta_{21}}\dot\alpha_{92}\ \mathbf{Com}\ \beta_{22}\right)\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }\frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \mathbf{(,)}\\&  : \left(\mathbf{Chan}\ \dot\alpha_{50}\xrightarrow{\beta_{25}}\left(\alpha_{24}\xrightarrow{\beta_{26}}\left(\mathbf{Chan}\ \dot\alpha_{50} \times \alpha_{24}\right)\right)\right)\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }\frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \mathbf{channel}\\&  : \left(\mathbf{()}\xrightarrow{\beta_{28}}\mathbf{Chan}\ \alpha_{27}\right)\\&  \& \left\{\left\{\mathbf{CHAN}\ \alpha_{27}\right\}\subseteq \beta_{28}\right\}\end{split}}}{\displaystyle \Downarrow }\frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \mathbf{()}\\&  : \mathbf{()}\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\mathbf{channel}\ \mathbf{()}\right)\\&  : \mathbf{Chan}\ \dot\alpha_{31}\\&  \& \left\{\left(\mathbf{()}\xrightarrow{\beta_{28}}\mathbf{Chan}\ \alpha_{27}\right)\subseteq \left(\mathbf{()}\xrightarrow{\beta_{30}}\mathbf{Chan}\ \dot\alpha_{31}\right), \left\{\mathbf{CHAN}\ \alpha_{27}\right\}\subseteq \beta_{28}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\left(\mathbf{()}\xrightarrow{\beta_{28}}\mathbf{Chan}\ \alpha_{27}\right)\subseteq \left(\mathbf{()}\xrightarrow{\beta_{30}}\mathbf{Chan}\ \dot\alpha_{31}\right), \left\{\mathbf{CHAN}\ \alpha_{27}\right\}\subseteq \beta_{28}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{27}\subseteq \dot\alpha_{31}, \dot\alpha_{31}\subseteq \alpha_{27}, \left\{\mathbf{CHAN}\ \alpha_{27}\right\}\subseteq \beta_{28}, \left\{\beta_{28}\right\}\subseteq \beta_{30}\right\}\\ \leadsto_{\mathcal{R}} & \mathbf{Chan}\ \dot\alpha_{31}\& \left\{\left\{\mathbf{CHAN}\ \dot\alpha_{31}\right\}\subseteq \beta_{28}, \left\{\beta_{28}\right\}\subseteq \beta_{30}\right\}\end{split}}}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\\&  : \left(\dot\alpha_{72}\xrightarrow{\dot\beta_{41}}\left(\mathbf{Chan}\ \dot\alpha_{51} \times \dot\alpha_{40}\right)\right)\\&  \& \left\{\left(\mathbf{Chan}\ \dot\alpha_{50}\xrightarrow{\beta_{25}}\left(\alpha_{24}\xrightarrow{\beta_{26}}\left(\mathbf{Chan}\ \dot\alpha_{50} \times \alpha_{24}\right)\right)\right)\subseteq \left(\mathbf{Chan}\ \dot\alpha_{31}\xrightarrow{\beta_{37}}\left(\dot\alpha_{72}\xrightarrow{\dot\beta_{41}}\left(\mathbf{Chan}\ \dot\alpha_{51} \times \dot\alpha_{40}\right)\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{31}\right\}\subseteq \beta_{28}, \left\{\beta_{28}\right\}\subseteq \beta_{30}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\left(\mathbf{Chan}\ \dot\alpha_{50}\xrightarrow{\beta_{25}}\left(\alpha_{24}\xrightarrow{\beta_{26}}\left(\mathbf{Chan}\ \dot\alpha_{50} \times \alpha_{24}\right)\right)\right)\subseteq \left(\mathbf{Chan}\ \dot\alpha_{31}\xrightarrow{\beta_{37}}\left(\dot\alpha_{72}\xrightarrow{\dot\beta_{41}}\left(\mathbf{Chan}\ \dot\alpha_{51} \times \dot\alpha_{40}\right)\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{31}\right\}\subseteq \beta_{28}, \left\{\beta_{28}\right\}\subseteq \beta_{30}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{24}\subseteq \dot\alpha_{40}, \dot\alpha_{31}\subseteq \dot\alpha_{50}, \dot\alpha_{50}\subseteq \dot\alpha_{31}, \dot\alpha_{50}\subseteq \dot\alpha_{51}, \dot\alpha_{51}\subseteq \dot\alpha_{50}, \dot\alpha_{72}\subseteq \alpha_{24}, \left\{\mathbf{CHAN}\ \dot\alpha_{31}\right\}\subseteq \beta_{28}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}\right\}\\ \leadsto_{\mathcal{R}} & \left(\dot\alpha_{72}\xrightarrow{\dot\beta_{41}}\left(\mathbf{Chan}\ \dot\alpha_{51} \times \dot\alpha_{72}\right)\right)\& \left\{\left\{\mathbf{CHAN}\ \dot\alpha_{51}\right\}\subseteq \beta_{28}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}\right\}\end{split}}\frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash y\\&  : \alpha_{2}\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\\&  : \left(\mathbf{Chan}\ \dot\alpha_{803} \times \dot\alpha_{76}\right)\\&  \& \left\{\left(\dot\alpha_{72}\xrightarrow{\dot\beta_{41}}\left(\mathbf{Chan}\ \dot\alpha_{51} \times \dot\alpha_{72}\right)\right)\subseteq \left(\alpha_{2}\xrightarrow{\beta_{71}}\left(\mathbf{Chan}\ \dot\alpha_{803} \times \dot\alpha_{76}\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{51}\right\}\subseteq \beta_{28}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\left(\dot\alpha_{72}\xrightarrow{\dot\beta_{41}}\left(\mathbf{Chan}\ \dot\alpha_{51} \times \dot\alpha_{72}\right)\right)\subseteq \left(\alpha_{2}\xrightarrow{\beta_{71}}\left(\mathbf{Chan}\ \dot\alpha_{803} \times \dot\alpha_{76}\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{51}\right\}\subseteq \beta_{28}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{2}\subseteq \dot\alpha_{72}, \dot\alpha_{51}\subseteq \dot\alpha_{803}, \dot\alpha_{72}\subseteq \dot\alpha_{76}, \dot\alpha_{803}\subseteq \dot\alpha_{51}, \left\{\mathbf{CHAN}\ \dot\alpha_{51}\right\}\subseteq \beta_{28}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{R}} & \left(\mathbf{Chan}\ \dot\alpha_{803} \times \alpha_{2}\right)\& \left\{\left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\\&  : \dot\alpha_{103}\ \mathbf{Com}\ \dot\beta_{104}\\&  \& \left\{\left(\left(\mathbf{Chan}\ \dot\alpha_{92} \times \dot\alpha_{92}\right)\xrightarrow{\beta_{21}}\dot\alpha_{92}\ \mathbf{Com}\ \beta_{22}\right)\subseteq \left(\left(\mathbf{Chan}\ \dot\alpha_{803} \times \alpha_{2}\right)\xrightarrow{\beta_{88}}\dot\alpha_{103}\ \mathbf{Com}\ \dot\beta_{104}\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\left(\left(\mathbf{Chan}\ \dot\alpha_{92} \times \dot\alpha_{92}\right)\xrightarrow{\beta_{21}}\dot\alpha_{92}\ \mathbf{Com}\ \beta_{22}\right)\subseteq \left(\left(\mathbf{Chan}\ \dot\alpha_{803} \times \alpha_{2}\right)\xrightarrow{\beta_{88}}\dot\alpha_{103}\ \mathbf{Com}\ \dot\beta_{104}\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{2}\subseteq \dot\alpha_{92}, \dot\alpha_{803}\subseteq \dot\alpha_{92}, \dot\alpha_{92}\subseteq \dot\alpha_{103}, \dot\alpha_{92}\subseteq \dot\alpha_{803}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{R}} & \dot\alpha_{803}\ \mathbf{Com}\ \dot\beta_{104}\& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\\&  : \dot\alpha_{132}\\&  \& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left(\dot\alpha_{131}\ \mathbf{Com}\ \beta_{19}\xrightarrow{\beta_{19}}\dot\alpha_{131}\right)\subseteq \left(\dot\alpha_{803}\ \mathbf{Com}\ \dot\beta_{104}\xrightarrow{\beta_{122}}\dot\alpha_{132}\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left(\dot\alpha_{131}\ \mathbf{Com}\ \beta_{19}\xrightarrow{\beta_{19}}\dot\alpha_{131}\right)\subseteq \left(\dot\alpha_{803}\ \mathbf{Com}\ \dot\beta_{104}\xrightarrow{\beta_{122}}\dot\alpha_{132}\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{131}\subseteq \dot\alpha_{132}, \dot\alpha_{803}\subseteq \dot\alpha_{131}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{R}} & \dot\alpha_{803}\& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\mathbf{(,)}\ \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\right)\\&  : \left(\dot\alpha_{181}\xrightarrow{\dot\beta_{160}}\left(\dot\alpha_{158} \times \dot\alpha_{159}\right)\right)\\&  \& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left(\dot\alpha_{150}\xrightarrow{\beta_{16}}\left(\alpha_{15}\xrightarrow{\beta_{17}}\left(\dot\alpha_{150} \times \alpha_{15}\right)\right)\right)\subseteq \left(\dot\alpha_{803}\xrightarrow{\beta_{146}}\left(\dot\alpha_{181}\xrightarrow{\dot\beta_{160}}\left(\dot\alpha_{158} \times \dot\alpha_{159}\right)\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left(\dot\alpha_{150}\xrightarrow{\beta_{16}}\left(\alpha_{15}\xrightarrow{\beta_{17}}\left(\dot\alpha_{150} \times \alpha_{15}\right)\right)\right)\subseteq \left(\dot\alpha_{803}\xrightarrow{\beta_{146}}\left(\dot\alpha_{181}\xrightarrow{\dot\beta_{160}}\left(\dot\alpha_{158} \times \dot\alpha_{159}\right)\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{15}\subseteq \dot\alpha_{159}, \alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{150}\subseteq \dot\alpha_{158}, \dot\alpha_{181}\subseteq \alpha_{15}, \dot\alpha_{803}\subseteq \dot\alpha_{150}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{R}} & \left(\dot\alpha_{181}\xrightarrow{\dot\beta_{160}}\left(\dot\alpha_{803} \times \dot\alpha_{181}\right)\right)\& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}\frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash x\\&  : \alpha_{10}\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\left(\mathbf{(,)}\ \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\right)\ x\right)\\&  : \left(\dot\alpha_{189} \times \dot\alpha_{190}\right)\\&  \& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left(\dot\alpha_{181}\xrightarrow{\dot\beta_{160}}\left(\dot\alpha_{803} \times \dot\alpha_{181}\right)\right)\subseteq \left(\alpha_{10}\xrightarrow{\beta_{180}}\left(\dot\alpha_{189} \times \dot\alpha_{190}\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left(\dot\alpha_{181}\xrightarrow{\dot\beta_{160}}\left(\dot\alpha_{803} \times \dot\alpha_{181}\right)\right)\subseteq \left(\alpha_{10}\xrightarrow{\beta_{180}}\left(\dot\alpha_{189} \times \dot\alpha_{190}\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{10}\subseteq \dot\alpha_{181}, \alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{181}\subseteq \dot\alpha_{190}, \dot\alpha_{803}\subseteq \dot\alpha_{189}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{R}} & \left(\dot\alpha_{803} \times \alpha_{10}\right)\& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\x& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{10}\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\right)\ x\right)\right)\\&  : \dot\alpha_{213}\\&  \& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left(\left(\dot\alpha_{222} \times \dot\alpha_{205}\right)\xrightarrow{\beta_{13}}\dot\alpha_{205}\right)\subseteq \left(\left(\dot\alpha_{803} \times \alpha_{10}\right)\xrightarrow{\beta_{204}}\dot\alpha_{213}\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left(\left(\dot\alpha_{222} \times \dot\alpha_{205}\right)\xrightarrow{\beta_{13}}\dot\alpha_{205}\right)\subseteq \left(\left(\dot\alpha_{803} \times \alpha_{10}\right)\xrightarrow{\beta_{204}}\dot\alpha_{213}\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{10}\subseteq \dot\alpha_{205}, \alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{205}\subseteq \dot\alpha_{213}, \dot\alpha_{803}\subseteq \dot\alpha_{222}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{R}} & \alpha_{10}\& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\lambda x.\ \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\right)\ x\right)\right)\right)\\&  : \left(\alpha_{10}\xrightarrow{\beta_{235}}\alpha_{10}\right)\\&  \& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}, \beta_{146}, \beta_{180}, \beta_{204}, \beta_{30}, \beta_{37}, \beta_{71}, \beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}, \beta_{146}, \beta_{180}, \beta_{204}, \beta_{30}, \beta_{37}, \beta_{71}, \beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{R}} & \left(\alpha_{10}\xrightarrow{\beta_{235}}\alpha_{10}\right)\& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \mathbf{if}\ \mathbf{True}\ \mathbf{then}\ f\ \mathbf{else}\ \left(\lambda x.\ \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\right)\ x\right)\right)\right)\\&  : \left(\dot\alpha_{250}\xrightarrow{\dot\beta_{253}}\dot\alpha_{891}\right)\\&  \& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \mathbf{Bool}\subseteq \mathbf{Bool}, \left(\alpha_{10}\xrightarrow{\beta_{235}}\alpha_{10}\right)\subseteq \left(\dot\alpha_{250}\xrightarrow{\dot\beta_{253}}\dot\alpha_{891}\right), \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\subseteq \left(\dot\alpha_{250}\xrightarrow{\dot\beta_{253}}\dot\alpha_{891}\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \mathbf{Bool}\subseteq \mathbf{Bool}, \left(\alpha_{10}\xrightarrow{\beta_{235}}\alpha_{10}\right)\subseteq \left(\dot\alpha_{250}\xrightarrow{\dot\beta_{253}}\dot\alpha_{891}\right), \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\subseteq \left(\dot\alpha_{250}\xrightarrow{\dot\beta_{253}}\dot\alpha_{891}\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{10}\subseteq \dot\alpha_{891}, \alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{250}\subseteq \alpha_{10}, \dot\alpha_{250}\subseteq \dot\alpha_{885}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{R}} & \left(\dot\alpha_{250}\xrightarrow{\dot\beta_{253}}\dot\alpha_{891}\right)\& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{250}\subseteq \dot\alpha_{885}, \dot\alpha_{250}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\mathbf{(,)}\ \mathbf{if}\ \mathbf{True}\ \mathbf{then}\ f\ \mathbf{else}\ \left(\lambda x.\ \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\right)\ x\right)\right)\right)\right)\\&  : \left(\dot\alpha_{406}\xrightarrow{\dot\beta_{310}}\left(\left(\dot\alpha_{335}\xrightarrow{\dot\beta_{338}}\dot\alpha_{336}\right) \times \dot\alpha_{309}\right)\right)\\&  \& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{250}\subseteq \dot\alpha_{885}, \dot\alpha_{250}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left(\left(\dot\alpha_{333}\xrightarrow{\dot\beta_{337}}\dot\alpha_{334}\right)\xrightarrow{\beta_{8}}\left(\alpha_{7}\xrightarrow{\beta_{9}}\left(\left(\dot\alpha_{333}\xrightarrow{\dot\beta_{337}}\dot\alpha_{334}\right) \times \alpha_{7}\right)\right)\right)\subseteq \left(\left(\dot\alpha_{250}\xrightarrow{\dot\beta_{253}}\dot\alpha_{891}\right)\xrightarrow{\beta_{277}}\left(\dot\alpha_{406}\xrightarrow{\dot\beta_{310}}\left(\left(\dot\alpha_{335}\xrightarrow{\dot\beta_{338}}\dot\alpha_{336}\right) \times \dot\alpha_{309}\right)\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{250}\subseteq \dot\alpha_{885}, \dot\alpha_{250}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left(\left(\dot\alpha_{333}\xrightarrow{\dot\beta_{337}}\dot\alpha_{334}\right)\xrightarrow{\beta_{8}}\left(\alpha_{7}\xrightarrow{\beta_{9}}\left(\left(\dot\alpha_{333}\xrightarrow{\dot\beta_{337}}\dot\alpha_{334}\right) \times \alpha_{7}\right)\right)\right)\subseteq \left(\left(\dot\alpha_{250}\xrightarrow{\dot\beta_{253}}\dot\alpha_{891}\right)\xrightarrow{\beta_{277}}\left(\dot\alpha_{406}\xrightarrow{\dot\beta_{310}}\left(\left(\dot\alpha_{335}\xrightarrow{\dot\beta_{338}}\dot\alpha_{336}\right) \times \dot\alpha_{309}\right)\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \alpha_{7}\subseteq \dot\alpha_{309}, \dot\alpha_{250}\subseteq \dot\alpha_{885}, \dot\alpha_{250}\subseteq \dot\alpha_{891}, \dot\alpha_{333}\subseteq \dot\alpha_{250}, \dot\alpha_{334}\subseteq \dot\alpha_{336}, \dot\alpha_{335}\subseteq \dot\alpha_{333}, \dot\alpha_{406}\subseteq \alpha_{7}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \dot\alpha_{891}\subseteq \dot\alpha_{334}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{R}} & \left(\dot\alpha_{406}\xrightarrow{\dot\beta_{310}}\left(\left(\dot\alpha_{335}\xrightarrow{\dot\beta_{338}}\dot\alpha_{891}\right) \times \dot\alpha_{406}\right)\right)\& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{335}\subseteq \dot\alpha_{885}, \dot\alpha_{335}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}\frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash y\\&  : \alpha_{2}\\&  \& \left\{\right\}\end{split}}}{\displaystyle \Downarrow }}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\left(\mathbf{(,)}\ \mathbf{if}\ \mathbf{True}\ \mathbf{then}\ f\ \mathbf{else}\ \left(\lambda x.\ \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\right)\ x\right)\right)\right)\right)\ y\right)\\&  : \left(\left(\dot\alpha_{451}\xrightarrow{\dot\beta_{454}}\dot\alpha_{452}\right) \times \dot\alpha_{453}\right)\\&  \& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{335}\subseteq \dot\alpha_{885}, \dot\alpha_{335}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left(\dot\alpha_{406}\xrightarrow{\dot\beta_{310}}\left(\left(\dot\alpha_{335}\xrightarrow{\dot\beta_{338}}\dot\alpha_{891}\right) \times \dot\alpha_{406}\right)\right)\subseteq \left(\alpha_{2}\xrightarrow{\beta_{404}}\left(\left(\dot\alpha_{451}\xrightarrow{\dot\beta_{454}}\dot\alpha_{452}\right) \times \dot\alpha_{453}\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{335}\subseteq \dot\alpha_{885}, \dot\alpha_{335}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left(\dot\alpha_{406}\xrightarrow{\dot\beta_{310}}\left(\left(\dot\alpha_{335}\xrightarrow{\dot\beta_{338}}\dot\alpha_{891}\right) \times \dot\alpha_{406}\right)\right)\subseteq \left(\alpha_{2}\xrightarrow{\beta_{404}}\left(\left(\dot\alpha_{451}\xrightarrow{\dot\beta_{454}}\dot\alpha_{452}\right) \times \dot\alpha_{453}\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{2}\subseteq \dot\alpha_{406}, \alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{335}\subseteq \dot\alpha_{885}, \dot\alpha_{335}\subseteq \dot\alpha_{891}, \dot\alpha_{406}\subseteq \dot\alpha_{453}, \dot\alpha_{451}\subseteq \dot\alpha_{335}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \dot\alpha_{891}\subseteq \dot\alpha_{452}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{R}} & \left(\left(\dot\alpha_{451}\xrightarrow{\dot\beta_{454}}\dot\alpha_{891}\right) \times \alpha_{2}\right)\& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{451}\subseteq \dot\alpha_{885}, \dot\alpha_{451}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\y& \mapsto \forall  . \left\{\right\} \Rightarrow \alpha_{2}\end{split}\right]\\&  \vdash \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \mathbf{if}\ \mathbf{True}\ \mathbf{then}\ f\ \mathbf{else}\ \left(\lambda x.\ \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\right)\ x\right)\right)\right)\right)\ y\right)\right)\\&  : \dot\alpha_{541}\\&  \& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{451}\subseteq \dot\alpha_{885}, \dot\alpha_{451}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left(\left(\left(\dot\alpha_{818}\xrightarrow{\dot\beta_{568}}\dot\alpha_{567}\right) \times \dot\alpha_{499}\right)\xrightarrow{\beta_{5}}\dot\alpha_{499}\right)\subseteq \left(\left(\left(\dot\alpha_{451}\xrightarrow{\dot\beta_{454}}\dot\alpha_{891}\right) \times \alpha_{2}\right)\xrightarrow{\beta_{498}}\dot\alpha_{541}\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{451}\subseteq \dot\alpha_{885}, \dot\alpha_{451}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left(\left(\left(\dot\alpha_{818}\xrightarrow{\dot\beta_{568}}\dot\alpha_{567}\right) \times \dot\alpha_{499}\right)\xrightarrow{\beta_{5}}\dot\alpha_{499}\right)\subseteq \left(\left(\left(\dot\alpha_{451}\xrightarrow{\dot\beta_{454}}\dot\alpha_{891}\right) \times \alpha_{2}\right)\xrightarrow{\beta_{498}}\dot\alpha_{541}\right), \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{2}\subseteq \dot\alpha_{499}, \alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{451}\subseteq \dot\alpha_{885}, \dot\alpha_{451}\subseteq \dot\alpha_{891}, \dot\alpha_{499}\subseteq \dot\alpha_{541}, \dot\alpha_{818}\subseteq \dot\alpha_{451}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \dot\alpha_{891}\subseteq \dot\alpha_{567}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}\right\}\\ \leadsto_{\mathcal{R}} & \alpha_{2}\& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}\right\}\end{split}}}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\end{split}\right]\\&  \vdash \left(\lambda y.\ \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \mathbf{if}\ \mathbf{True}\ \mathbf{then}\ f\ \mathbf{else}\ \left(\lambda x.\ \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\right)\ x\right)\right)\right)\right)\ y\right)\right)\right)\\&  : \left(\alpha_{2}\xrightarrow{\beta_{611}}\alpha_{2}\right)\\&  \& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}, \beta_{404}, \beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}, \beta_{404}, \beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}\right\}\\ \leadsto_{\mathcal{R}} & \left(\alpha_{2}\xrightarrow{\beta_{611}}\alpha_{2}\right)\& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}\right\}\end{split}}\frac{\displaystyle \frac{\displaystyle \frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\id& \mapsto \forall \alpha_{2}\dot\alpha_{612}\beta_{122}\beta_{13}\beta_{146}\beta_{16}\beta_{17}\beta_{180}\beta_{19}\beta_{204}\beta_{21}\beta_{22}\beta_{235}\beta_{25}\beta_{26}\beta_{277}\beta_{28}\beta_{30}\beta_{37}\beta_{404}\beta_{498}\beta_{5}\beta_{611}\beta_{71}\beta_{88}\beta_{8}\beta_{9}\dot\beta_{104}\dot\beta_{160}\dot\beta_{253}\dot\beta_{310}\dot\beta_{337}\dot\beta_{338}\dot\beta_{41}\dot\beta_{454}\dot\beta_{568} . \left\{\alpha_{2}\subseteq \dot\alpha_{612}, \left\{\mathbf{CHAN}\ \dot\alpha_{612}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}\right\} \Rightarrow \left(\alpha_{2}\xrightarrow{\beta_{611}}\alpha_{2}\right)\end{split}\right]\\&  \vdash id\\&  : \left(\left(\dot\alpha_{880}\xrightarrow{\dot\beta_{744}}\dot\alpha_{739}\right)\xrightarrow{\beta_{663}}\left(\dot\alpha_{880}\xrightarrow{\dot\beta_{744}}\dot\alpha_{739}\right)\right)\\&  \& \left\{\left(\dot\alpha_{880}\xrightarrow{\dot\beta_{744}}\dot\alpha_{739}\right)\subseteq \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right), \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}\right\}\end{split}}}{\displaystyle \Downarrow }\frac{\displaystyle \frac{\displaystyle }{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\id& \mapsto \forall \alpha_{2}\dot\alpha_{612}\beta_{122}\beta_{13}\beta_{146}\beta_{16}\beta_{17}\beta_{180}\beta_{19}\beta_{204}\beta_{21}\beta_{22}\beta_{235}\beta_{25}\beta_{26}\beta_{277}\beta_{28}\beta_{30}\beta_{37}\beta_{404}\beta_{498}\beta_{5}\beta_{611}\beta_{71}\beta_{88}\beta_{8}\beta_{9}\dot\beta_{104}\dot\beta_{160}\dot\beta_{253}\dot\beta_{310}\dot\beta_{337}\dot\beta_{338}\dot\beta_{41}\dot\beta_{454}\dot\beta_{568} . \left\{\alpha_{2}\subseteq \dot\alpha_{612}, \left\{\mathbf{CHAN}\ \dot\alpha_{612}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}\right\} \Rightarrow \left(\alpha_{2}\xrightarrow{\beta_{611}}\alpha_{2}\right)\end{split}\right]\\&  \vdash id\\&  : \left(\alpha_{680}\xrightarrow{\beta_{702}}\alpha_{680}\right)\\&  \& \left\{\alpha_{680}\subseteq \dot\alpha_{902}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}\right\}\end{split}}}{\displaystyle \Downarrow }}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\\id& \mapsto \forall \alpha_{2}\dot\alpha_{612}\beta_{122}\beta_{13}\beta_{146}\beta_{16}\beta_{17}\beta_{180}\beta_{19}\beta_{204}\beta_{21}\beta_{22}\beta_{235}\beta_{25}\beta_{26}\beta_{277}\beta_{28}\beta_{30}\beta_{37}\beta_{404}\beta_{498}\beta_{5}\beta_{611}\beta_{71}\beta_{88}\beta_{8}\beta_{9}\dot\beta_{104}\dot\beta_{160}\dot\beta_{253}\dot\beta_{310}\dot\beta_{337}\dot\beta_{338}\dot\beta_{41}\dot\beta_{454}\dot\beta_{568} . \left\{\alpha_{2}\subseteq \dot\alpha_{612}, \left\{\mathbf{CHAN}\ \dot\alpha_{612}\right\}\subseteq \beta_{28}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}\right\} \Rightarrow \left(\alpha_{2}\xrightarrow{\beta_{611}}\alpha_{2}\right)\end{split}\right]\\&  \vdash \left(id\ id\right)\\&  : \left(\dot\alpha_{740}\xrightarrow{\dot\beta_{745}}\dot\alpha_{741}\right)\\&  \& \left\{\alpha_{680}\subseteq \dot\alpha_{902}, \left(\dot\alpha_{880}\xrightarrow{\dot\beta_{744}}\dot\alpha_{739}\right)\subseteq \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right), \left(\left(\dot\alpha_{880}\xrightarrow{\dot\beta_{744}}\dot\alpha_{739}\right)\xrightarrow{\beta_{663}}\left(\dot\alpha_{880}\xrightarrow{\dot\beta_{744}}\dot\alpha_{739}\right)\right)\subseteq \left(\left(\alpha_{680}\xrightarrow{\beta_{702}}\alpha_{680}\right)\xrightarrow{\beta_{720}}\left(\dot\alpha_{740}\xrightarrow{\dot\beta_{745}}\dot\alpha_{741}\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{680}\subseteq \dot\alpha_{902}, \left(\dot\alpha_{880}\xrightarrow{\dot\beta_{744}}\dot\alpha_{739}\right)\subseteq \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right), \left(\left(\dot\alpha_{880}\xrightarrow{\dot\beta_{744}}\dot\alpha_{739}\right)\xrightarrow{\beta_{663}}\left(\dot\alpha_{880}\xrightarrow{\dot\beta_{744}}\dot\alpha_{739}\right)\right)\subseteq \left(\left(\alpha_{680}\xrightarrow{\beta_{702}}\alpha_{680}\right)\xrightarrow{\beta_{720}}\left(\dot\alpha_{740}\xrightarrow{\dot\beta_{745}}\dot\alpha_{741}\right)\right), \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{680}\subseteq \dot\alpha_{739}, \alpha_{680}\subseteq \dot\alpha_{902}, \dot\alpha_{739}\subseteq \dot\alpha_{741}, \dot\alpha_{739}\subseteq \dot\alpha_{914}, \dot\alpha_{740}\subseteq \dot\alpha_{880}, \dot\alpha_{742}\subseteq \dot\alpha_{880}, \dot\alpha_{880}\subseteq \alpha_{680}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{663}\right\}\subseteq \beta_{720}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{702}\right\}\subseteq \dot\beta_{744}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{745}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{746}\right\}\\ \leadsto_{\mathcal{R}} & \left(\dot\alpha_{880}\xrightarrow{\dot\beta_{745}}\dot\alpha_{880}\right)\& \left\{\dot\alpha_{742}\subseteq \dot\alpha_{880}, \dot\alpha_{880}\subseteq \dot\alpha_{902}, \dot\alpha_{880}\subseteq \dot\alpha_{914}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{663}\right\}\subseteq \beta_{720}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{702}\right\}\subseteq \dot\beta_{744}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{745}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{746}\right\}\end{split}}}{\displaystyle \begin{split}& \left[\begin{split}f& \mapsto \forall  . \left\{\right\} \Rightarrow \left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\end{split}\right]\\&  \vdash \mathbf{let}\ id\ \mathbf{=}\ \left(\lambda y.\ \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \mathbf{if}\ \mathbf{True}\ \mathbf{then}\ f\ \mathbf{else}\ \left(\lambda x.\ \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\right)\ x\right)\right)\right)\right)\ y\right)\right)\right)\ \mathbf{in}\ \left(id\ id\right)\\&  : \left(\dot\alpha_{880}\xrightarrow{\dot\beta_{745}}\dot\alpha_{880}\right)\\&  \& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{742}\subseteq \dot\alpha_{880}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{880}\subseteq \dot\alpha_{902}, \dot\alpha_{880}\subseteq \dot\alpha_{914}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{663}\right\}\subseteq \beta_{720}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{702}\right\}\subseteq \dot\beta_{744}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{745}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{746}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{742}\subseteq \dot\alpha_{880}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{880}\subseteq \dot\alpha_{902}, \dot\alpha_{880}\subseteq \dot\alpha_{914}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{663}\right\}\subseteq \beta_{720}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{702}\right\}\subseteq \dot\beta_{744}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{745}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{746}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\alpha_{2}\subseteq \dot\alpha_{803}, \dot\alpha_{742}\subseteq \dot\alpha_{880}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{880}\subseteq \dot\alpha_{902}, \dot\alpha_{880}\subseteq \dot\alpha_{914}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{663}\right\}\subseteq \beta_{720}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{702}\right\}\subseteq \dot\beta_{744}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{745}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{746}\right\}\\ \leadsto_{\mathcal{R}} & \left(\dot\alpha_{880}\xrightarrow{\dot\beta_{745}}\dot\alpha_{880}\right)\& \left\{\dot\alpha_{742}\subseteq \dot\alpha_{880}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{880}\subseteq \dot\alpha_{902}, \dot\alpha_{880}\subseteq \dot\alpha_{914}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{663}\right\}\subseteq \beta_{720}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{702}\right\}\subseteq \dot\beta_{744}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{745}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{746}\right\}\end{split}}}{\displaystyle \begin{split}& \epsilon \\&  \vdash \left(\lambda f.\ \mathbf{let}\ id\ \mathbf{=}\ \left(\lambda y.\ \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \mathbf{if}\ \mathbf{True}\ \mathbf{then}\ f\ \mathbf{else}\ \left(\lambda x.\ \left(\mathbf{snd}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{sync}\ \left(\mathbf{send}\ \left(\left(\mathbf{(,)}\ \left(\mathbf{channel}\ \mathbf{()}\right)\right)\ y\right)\right)\right)\right)\ x\right)\right)\right)\right)\ y\right)\right)\right)\ \mathbf{in}\ \left(id\ id\right)\right)\\&  : \left(\left(\dot\alpha_{885}\xrightarrow{\dot\beta_{252}}\dot\alpha_{890}\right)\xrightarrow{\beta_{876}}\left(\dot\alpha_{880}\xrightarrow{\dot\beta_{745}}\dot\alpha_{880}\right)\right)\\&  \& \left\{\dot\alpha_{742}\subseteq \dot\alpha_{880}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{880}\subseteq \dot\alpha_{902}, \dot\alpha_{880}\subseteq \dot\alpha_{914}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{663}\right\}\subseteq \beta_{720}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{702}\right\}\subseteq \dot\beta_{744}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{720}\right\}\subseteq \beta_{876}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{745}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{746}\right\}\end{split}}}{\displaystyle \begin{split}& \left\{\dot\alpha_{742}\subseteq \dot\alpha_{880}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{880}\subseteq \dot\alpha_{902}, \dot\alpha_{880}\subseteq \dot\alpha_{914}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{663}\right\}\subseteq \beta_{720}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{702}\right\}\subseteq \dot\beta_{744}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{720}\right\}\subseteq \beta_{876}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{745}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{746}\right\}\\ \leadsto_{\mathcal{F}} & \left\{\dot\alpha_{742}\subseteq \dot\alpha_{880}, \dot\alpha_{818}\subseteq \dot\alpha_{885}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{880}\subseteq \dot\alpha_{902}, \dot\alpha_{880}\subseteq \dot\alpha_{914}, \dot\alpha_{890}\subseteq \dot\alpha_{891}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{663}\right\}\subseteq \beta_{720}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{702}\right\}\subseteq \dot\beta_{744}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{720}\right\}\subseteq \beta_{876}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{745}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{746}\right\}\\ \leadsto_{\mathcal{R}} & \left(\left(\dot\alpha_{818}\xrightarrow{\dot\beta_{252}}\dot\alpha_{891}\right)\xrightarrow{\beta_{876}}\left(\dot\alpha_{880}\xrightarrow{\dot\beta_{745}}\dot\alpha_{880}\right)\right)\& \left\{\dot\alpha_{742}\subseteq \dot\alpha_{880}, \dot\alpha_{818}\subseteq \dot\alpha_{891}, \dot\alpha_{880}\subseteq \dot\alpha_{902}, \dot\alpha_{880}\subseteq \dot\alpha_{914}, \left\{\mathbf{CHAN}\ \dot\alpha_{803}\right\}\subseteq \beta_{28}, \left\{\mathbf{CHAN}\ \dot\alpha_{902}\right\}\subseteq \beta_{696}, \left\{\mathbf{CHAN}\ \left(\dot\alpha_{742}\xrightarrow{\dot\beta_{746}}\dot\alpha_{914}\right)\right\}\subseteq \beta_{657}, \left\{\beta_{122}\right\}\subseteq \beta_{235}, \left\{\beta_{13}\right\}\subseteq \beta_{204}, \left\{\beta_{146}\right\}\subseteq \beta_{235}, \left\{\beta_{16}\right\}\subseteq \beta_{146}, \left\{\beta_{17}\right\}\subseteq \dot\beta_{160}, \left\{\beta_{180}\right\}\subseteq \beta_{235}, \left\{\beta_{19}\right\}\subseteq \beta_{122}, \left\{\beta_{204}\right\}\subseteq \beta_{235}, \left\{\beta_{21}\right\}\subseteq \beta_{88}, \left\{\beta_{22}\right\}\subseteq \dot\beta_{104}, \left\{\beta_{235}\right\}\subseteq \dot\beta_{253}, \left\{\beta_{25}\right\}\subseteq \beta_{37}, \left\{\beta_{26}\right\}\subseteq \dot\beta_{41}, \left\{\beta_{277}\right\}\subseteq \beta_{611}, \left\{\beta_{28}\right\}\subseteq \beta_{30}, \left\{\beta_{30}\right\}\subseteq \beta_{235}, \left\{\beta_{37}\right\}\subseteq \beta_{235}, \left\{\beta_{404}\right\}\subseteq \beta_{611}, \left\{\beta_{498}\right\}\subseteq \beta_{611}, \left\{\beta_{5}\right\}\subseteq \beta_{498}, \left\{\beta_{643}\right\}\subseteq \beta_{653}, \left\{\beta_{644}\right\}\subseteq \beta_{650}, \left\{\beta_{645}\right\}\subseteq \beta_{653}, \left\{\beta_{646}\right\}\subseteq \beta_{645}, \left\{\beta_{647}\right\}\subseteq \beta_{669}, \left\{\beta_{648}\right\}\subseteq \beta_{653}, \left\{\beta_{649}\right\}\subseteq \beta_{643}, \left\{\beta_{650}\right\}\subseteq \beta_{653}, \left\{\beta_{651}\right\}\subseteq \beta_{665}, \left\{\beta_{652}\right\}\subseteq \beta_{668}, \left\{\beta_{653}\right\}\subseteq \beta_{670}, \left\{\beta_{654}\right\}\subseteq \beta_{659}, \left\{\beta_{655}\right\}\subseteq \beta_{674}, \left\{\beta_{656}\right\}\subseteq \beta_{663}, \left\{\beta_{657}\right\}\subseteq \beta_{658}, \left\{\beta_{658}\right\}\subseteq \beta_{653}, \left\{\beta_{659}\right\}\subseteq \beta_{653}, \left\{\beta_{660}\right\}\subseteq \beta_{663}, \left\{\beta_{661}\right\}\subseteq \beta_{663}, \left\{\beta_{662}\right\}\subseteq \beta_{661}, \left\{\beta_{663}\right\}\subseteq \beta_{720}, \left\{\beta_{664}\right\}\subseteq \beta_{653}, \left\{\beta_{665}\right\}\subseteq \beta_{653}, \left\{\beta_{666}\right\}\subseteq \beta_{656}, \left\{\beta_{667}\right\}\subseteq \beta_{671}, \left\{\beta_{668}\right\}\subseteq \beta_{649}, \left\{\beta_{669}\right\}\subseteq \beta_{648}, \left\{\beta_{670}\right\}\subseteq \beta_{672}, \left\{\beta_{671}\right\}\subseteq \beta_{660}, \left\{\beta_{672}\right\}\subseteq \beta_{673}, \left\{\beta_{673}\right\}\subseteq \beta_{675}, \left\{\beta_{674}\right\}\subseteq \beta_{664}, \left\{\beta_{675}\right\}\subseteq \beta_{676}, \left\{\beta_{682}\right\}\subseteq \beta_{692}, \left\{\beta_{683}\right\}\subseteq \beta_{689}, \left\{\beta_{684}\right\}\subseteq \beta_{692}, \left\{\beta_{685}\right\}\subseteq \beta_{684}, \left\{\beta_{686}\right\}\subseteq \beta_{708}, \left\{\beta_{687}\right\}\subseteq \beta_{692}, \left\{\beta_{688}\right\}\subseteq \beta_{682}, \left\{\beta_{689}\right\}\subseteq \beta_{692}, \left\{\beta_{690}\right\}\subseteq \beta_{704}, \left\{\beta_{691}\right\}\subseteq \beta_{707}, \left\{\beta_{692}\right\}\subseteq \beta_{709}, \left\{\beta_{693}\right\}\subseteq \beta_{698}, \left\{\beta_{694}\right\}\subseteq \beta_{713}, \left\{\beta_{695}\right\}\subseteq \beta_{702}, \left\{\beta_{696}\right\}\subseteq \beta_{697}, \left\{\beta_{697}\right\}\subseteq \beta_{692}, \left\{\beta_{698}\right\}\subseteq \beta_{692}, \left\{\beta_{699}\right\}\subseteq \beta_{702}, \left\{\beta_{700}\right\}\subseteq \beta_{702}, \left\{\beta_{701}\right\}\subseteq \beta_{700}, \left\{\beta_{702}\right\}\subseteq \dot\beta_{744}, \left\{\beta_{703}\right\}\subseteq \beta_{692}, \left\{\beta_{704}\right\}\subseteq \beta_{692}, \left\{\beta_{705}\right\}\subseteq \beta_{695}, \left\{\beta_{706}\right\}\subseteq \beta_{710}, \left\{\beta_{707}\right\}\subseteq \beta_{688}, \left\{\beta_{708}\right\}\subseteq \beta_{687}, \left\{\beta_{709}\right\}\subseteq \beta_{711}, \left\{\beta_{710}\right\}\subseteq \beta_{699}, \left\{\beta_{711}\right\}\subseteq \beta_{712}, \left\{\beta_{712}\right\}\subseteq \beta_{714}, \left\{\beta_{713}\right\}\subseteq \beta_{703}, \left\{\beta_{714}\right\}\subseteq \beta_{715}, \left\{\beta_{71}\right\}\subseteq \beta_{235}, \left\{\beta_{720}\right\}\subseteq \beta_{876}, \left\{\beta_{88}\right\}\subseteq \beta_{235}, \left\{\beta_{8}\right\}\subseteq \beta_{277}, \left\{\beta_{9}\right\}\subseteq \dot\beta_{310}, \left\{\dot\beta_{104}\right\}\subseteq \beta_{19}, \left\{\dot\beta_{160}\right\}\subseteq \beta_{180}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{670}, \left\{\dot\beta_{252}\right\}\subseteq \beta_{709}, \left\{\dot\beta_{252}\right\}\subseteq \dot\beta_{253}, \left\{\dot\beta_{253}\right\}\subseteq \dot\beta_{337}, \left\{\dot\beta_{310}\right\}\subseteq \beta_{404}, \left\{\dot\beta_{337}\right\}\subseteq \dot\beta_{338}, \left\{\dot\beta_{338}\right\}\subseteq \dot\beta_{454}, \left\{\dot\beta_{41}\right\}\subseteq \beta_{71}, \left\{\dot\beta_{454}\right\}\subseteq \dot\beta_{568}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{745}, \left\{\dot\beta_{744}\right\}\subseteq \dot\beta_{746}\right\}\end{split}}
\end{gather}
\end{document}