### 2015-09-01 deprecated/pattern-match-analysis/thesis/abstract.tex

\chapter*{\centering \begin{normalsize}Abstract\end{normalsize}}
\begin{quotation}
\noindent
In this thesis we attempt to answer the research question:
\begin{quote}
Can we use a type and effect system in combination with refinement types to develop a pattern-match analysis for a non-strict higher-order functional language that is both performant and precise enough to be of practical use?
\end{quote}

In Chapter \ref{chapintroduction} we present a number of examples to demonstrate why this is an interesting problem. In Chapter \ref{chapcontext} we give a short introduction to the relevant concepts of the research question: higher-order functional languages and type and effect systems. In Chapter \ref{chappma} we give an overview of the pattern match analysis we developed and give a detailed description of the constraint generation and constraint solving phases in respectively Chapter \ref{chapcg} and Chapter \ref{chapcs}. In Chapter \ref{chapimplementation} we discussion the implementation of the analysis we have built. We evaluate the effectiveness and limitations of our analysis in Chapter \ref{chapevaluation}. In Chapter \ref{chaprelatedwork} we present work related to our research question and discuss which aspects of that work are relevant to or different from our proposed system. Finally, in Chapter \ref{chapfurtherresearch}, we propose a number of directions for further research and future implementation work to improve the precision and applicability of the analysis.

\end{quotation}
\clearpage