### 2015-09-01 deprecated/pattern-match-analysis/thesis/implementation.tex

\chapter{Implementation}\label{chapimplementation}

An implementation of the analysis has been implemented in Haskell. It has been built on top of the \texttt{haskell-src-exts} library, which is used for both parsing of the input and representation of the abstract syntax tree.

Refinements and other extensions on Haskell's type system that are required for the analysis are represented using the annotation fields available in the abstract syntax of \texttt{Language.Haskell.Exts.Annotated}.

An overview of the modules that comprise the implementation can be found in Figure \ref{moduleoverview}.

Additionally, we have build a collection of example functions that where used for regression testing purposes during the development of the analysis and its implementation.

A demonstration of the implementation can be found in Appendix \ref{appx}.

\vspace{5em}

\begin{figure}[h]
\centering
\begin{tabular}{lrl}
Module & Lines & Description \\
\hline
\texttt{Main}          &  134 & Main \\
\texttt{TypeInference} &  391 & Constraint generation \\
\texttt{PatternTyping} &  153 & Typing of patterns \\
\texttt{Solver}        &  263 & Constraint solving \\
\texttt{Common}        &  342 & Substitution and unification \\
\texttt{Refinements}   &  152 & Refinements (representation and operations) \\
\texttt{Abstract.*}    &  355 & Abstract values (unit, booleans, integers, lists)\\
\texttt{Util.*}        &   48 & Miscellaneous
\end{tabular}
\caption{Module overview}
\label{moduleoverview}
\end{figure}